---
title: "The Weathers Effect on Rooftop Bar Revenues"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE
                      , show = TRUE
                      , fig.width=10)
```

We're interested in the revenues of a rooftop bar in NYC vs the weather on corresponding days. Is there and headroom for revenue growth by, for example, building an enclosure? Would heating or covering the roofdeck bring in additional revenue? How much do you lose per year to rain?


# Data

The revenue data was obtained independently directly from the hotel bar. The weather data was scraped from [weatherunderground.com](https://www.wunderground.com/) using the `library(weatherData)` package. 

There are many potenital "stations" you can gather weather data from. It is important to find not just the most physically proximate station, but also to find a station with consitent, clean data. Some of the data API functions did not return the correct data, or the time series it returned was unduly censcored. to that end, `getStationCode("New York", region = "NY")` was helpful in retrieving the most reliable station for the NYC area: LGA. 

I modified the `getWeatherForYear` function slightly, as the out-of-the-box solution wasn't scraping the API correctly. In addition, I found that scraping the years individually worked, while supplying a range of dates did not. If you wanted to scrape several years of data, one of the `purrr::map()` funtions could make that possible.  


```{r, echo=TRUE, message=FALSE, warning=FALSE, show=T}

getWeatherForYear2 <- 
  function (station_id, year, station_type = "airportCode", opt_detailed = TRUE, 
            opt_write_to_file = FALSE, opt_all_columns = TRUE) 
  {
    if (year == (1900 + as.POSIXlt(Sys.Date())$year)) {
      last_day <- Sys.Date()
    }
    else {
      last_day <- paste0(year, "-12-31")
    }
    first_day <- paste0(year, "-01-01")
    getWeatherForDate(station_id, first_day, last_day, station_type, 
                      opt_detailed, opt_all_columns = opt_all_columns)
  }

```


The final is was concatenated into a single dataframe and written to disk for analysis:

```{r}

suppressMessages({
  
  library(tidyverse)
  readr::read_csv("Weather Data/Weather-NYC-2017-05-30.csv", n_max = 5) %>% knitr::kable() 
  
})
```


# EDA

For a complete list of EDA plots, run the `03-weather-revenue-EDA.R` script. Amoung many interesting relationships, the correlation betwen Temperature and Earnings proved visually appealing. Various nightly specials (many of which could be considered outliers), also proved to be an interesting candidate for the modeling step:


```{r, echo=TRUE, fig.width=10}
library(tidyverse)
rev_ts <- read_rds("weather-revenue-data-v002.rds")

rev_ts_model <- 
  rev_ts %>% 
  mutate(Special_ind = if_else(is.na(Special),FALSE,TRUE)) %>% 
  filter(Earnings>0) %>% 
  filter(!is.na(Earnings))


# specials indicator ------------------------------------------------------
rev_ts_model %>% ggplot() + 
  aes(x = Earnings, y = Mean_TemperatureF, label = Special) + 
  geom_point(aes(color = Special_ind)) + 
  geom_smooth(method="lm")+
  theme_minimal()+
  ggthemes::scale_color_fivethirtyeight()+
  labs(title = "Do Specials Earn Extra Money?"
       ,y = "Mean Tempurature (F)"
       ,col = "Special?")

```



# Modeling

Inference is our goal here (rather than predictive capability). For this case, a simple linear model may prove to be the best tool since they are comuputationally negligible and hihgly interpretable.

Let's look at the relationship bettwen `Earnings` and both `Temp` and `Precipitation`

```{r}

library(tidyverse)

rev_ts <- read_rds("weather-revenue-data-v002.rds")

rev_ts_model <- 
  rev_ts %>% mutate(Happy_Hour = grepl("Happy Hour",Special)
                    , Special_ind = ifelse(is.na(Special),FALSE,TRUE)
                    , Rain_ind = ifelse(PrecipitationIn>0.5,TRUE,FALSE)
                    , Rain_ind = ifelse(is.na(Rain_ind),FALSE,Rain_ind)
                    , Events = factor(Events)) %>% 
  filter(Earnings>0) %>% 
  filter(!is.na(Earnings))


# modeling temp -----------------------------------------------------------

f4 <- as.formula(Earnings ~ Mean_TemperatureF)

f4_lm <- lm(formula = f4
            , data = rev_ts_model
)
summary(f4_lm)
```

With an adjusted R squared of only `r round(as.numeric(summary(f4_lm)$adj.r.squared),2)`, the model accounts for very little of the error in the data. From our EDA above, we can conclude that the relationship between `Temperature` and `Earnings` is not strictly linear. In fact, it appears that when tempuratures are too hot, the Earnings go down somewhat, creating a polynomial distribution similar to a parabola, or a rainbow. 

Fitting a polynomial model to the data reveals that earnings peak around 75 degrees F, then start to decline. This suggests that days that are too hot are costing the bar money.

```{r}
library(modelr)


f4_poly <- as.formula(Earnings ~ poly(Mean_TemperatureF,5))

f4_poly_lm <- lm(formula = f4_poly
                 , data = rev_ts_model %>% filter(Earnings<4000)
)

rev_ts_model %>% 
  data_grid(Mean_TemperatureF) %>% 
  add_predictions(f4_poly_lm) %>% 
  ggplot()+
  aes(x = Mean_TemperatureF, y  = pred) +
  geom_line(color="skyblue", size = 2)+
  geom_point(data=rev_ts_model %>% filter(Earnings<4000), aes(x = Mean_TemperatureF, y = Earnings))+
  geom_vline(aes(xintercept = 75),color = "darkgreen", alpha = 0.5)+
  theme_minimal()+
  ggthemes::scale_color_tableau()+
  labs(y = "Earnings"
       ,x = "Mean Temperature")

summary(f4_poly_lm)
```


Rain, on the other hand, seems to have little to no linear relation to Earnings. This is due to the fact that many days have very little precipitation, which may occur in the morning or afternoon (when the bar isn't open). A more robust feature here would be precipitation during operating hours. This could be scraped from the wearther data API using `detailed=TRUE` in the call, however, the incremental benefit this would bring us is not worth the effort for this explorartory exercize. 


```{r}

library(modelr)

# effect of rain on response ----------------------------------------------

f5 <- as.formula(Earnings ~ PrecipitationIn)

f5_lm <- lm(formula = f5
            , data = rev_ts_model
)

rev_ts_model %>% 
  filter(PrecipitationIn>0) %>% 
  data_grid(PrecipitationIn) %>% 
  add_predictions(f5_lm) %>% 
  ggplot()+
  aes(x = PrecipitationIn, y  = pred) +
  geom_line(color="skyblue", size = 2)+
  geom_point(data=rev_ts_model %>% filter(Earnings<4000, PrecipitationIn>0), aes(x = PrecipitationIn, y = Earnings))+
  #geom_vline(aes(xintercept = 75),color = "darkgreen", alpha = 0.5)+
  theme_minimal()+
  ggthemes::scale_color_tableau()+
  labs(y = "Earnings"
       ,x = "Precipitation in inches")

summary(f4_poly_lm)

summary(f5_lm)

```


Finally, let's look at whether advertising a special increases the Earings in a meaningful way

```{r}
rev_ts_model %>% 
  ggplot()+
  aes(x = Special_ind, y = Earnings, color = Special_ind, group = Special_ind)+
  geom_boxplot()+
  geom_jitter()+
  ggthemes::scale_color_fivethirtyeight()+
  theme_minimal()+
  theme(legend.position = "none")+
  labs(x = "Was a special advertised?"
       ,col = NULL)

```

Earnings seem to increase slightly when offers (e.g., Hapy Hour specials) are offered. Is this statisticaly significant? According to our model, there is a weak but statistically significant linear relationship between Specials and Earnings. The model estimates that, all else equal, Specials increase the daily earnings by about $200 on average.

```{r}
# effect of advertising a speal on response -------------------------------

f6 <- as.formula(
  Earnings ~ Special_ind
  )

f6_lm <- lm(formula = f6, data = rev_ts_model)
summary(f6_lm)

```

# Conclusion
Temperature can be used to effectively model Rooftop Bar Earnings up to a point. The relationship is not linear, however. Earnings tend to peak around 75 degrees (F), then trail off as the air gets warmer. Reccomendations to increase revenue might include building an enclosure of some kind to provide both heating and shade which will enable the bar to operate more days out of the year near full capacity. 








